---
title: "Cheung_affinity"
author: "Dvora"
date: "19 12 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---
This function is getting Cheung Temperature affinity. He create table of 991 species from the marine ecosystem (not fish only). 

Chengy paper:

Signature of ocean warming in global fisheries catch (2013)
https://www.nature.com/articles/nature12156

input:"species" - species name, vector or dataframe, could recieved 1 species name or a vector/data frame with a lot of species names.

output:
species_tp - data frame with TP50, TP75, TP25 - temperature preference of the species, if it exist in Chueng table.
If the species doesn't exist, recieved NA

---!!!NOTICE!!!---

If you don't accept temperatue, try synonym names - Chueng table is from 2013...

```{r}
library(tidyverse)

```



```{r}
#upload Chueng data
ch_data = read.csv(file = "Chueng data clean.csv")

#replace the name of species column, and remove coomon English names
ch_data = ch_data %>% rename(species = Latin)%>% select(-English)

#convert TP columns to numeric
ch_data$TP25 = as.numeric(ch_data$TP25)
ch_data$TP75 = as.numeric(ch_data$TP75)

#left join
species_tp  = as.data.frame(species)
species_tp =species_tp %>% left_join(ch_data)


```

