---
title: "GrossAff"
author: "Dvora"
date: "18 12 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---


GrossAff - Gross affinity from GBIF data 

input – overlay GetCoords + GetEnv
   
  1.  rasters - 'RasterStack' file, with layers of temperature. The choosen layres writen in "names" raw.(e.g BO_sstmax, BO_sstmean )

  2. data frame with species names, lat, lon

output: summary statistics (mean, median, mode, quantiles, SD, range, etc.)
  "stat_dataframe" - data frame withe the summaries values, any row is temerature layer (mean SST, max SST inc. - the layers chooden at previeos functions)




Load packages
```{r}
library(summarytools)
library(tidyverse)
library(sdmpredictors)
library(dismo)
library(jsonlite)


```

cerate example data, with coordinate for Boops boops.
the data coords has the coordinate, with "lat" and "lon" column
`coords` should be a dataframe of coordinates (lat and lon)


```{r}
sp_data = read.csv(file = "sp.data.CSV")
  sp_data = sp_data %>% filter(species=="Boops boops")
    coords = sp_data %>% dplyr:: select(lon,lat)

```


overlap the coordinates and the temperatures layers from the raster
and create tible withe the summary for any layer

```{r}
#lapplay loop for overlap and summary
list_of_layer_stats <- lapply(names(rasters),function(layer_name)
  
  {
  # extract according to 'coords'
    temp <- raster::extract(rasters[[layer_name]], coords, df = TRUE) 
      sumdata <- summarytools::descr(temp)
      sumdata = as_tibble(t(sumdata))[1,]
  })

#join the names of the layers and convert the list of summaries to 1 data frame, "stat_dataframe"
  names(list_of_layer_stats) <- names(rasters)
    stat_dataframe <- bind_rows(list_of_layer_stats,.id = "id") %>%   rename(layer = id) %>% 
      mutate(Range = Max-Min) #add range collumn
    
```


