---
title: "GrossAff"
author: "Dvora"
date: "18 12 2019"
output: html_document
---
GetCoords: Download data from GBIF and OBIS  -  Ori and Sarah
input: species list
output: coordinates dataframe


GetEnv: Load environmental data (e.g. temperature) from Bio-ORACLE and GMED  - | Ori and Sarah
input: Define a resolution and extent.
Load environmental layers


GrossAff - Gross affinity from GBIF data 

input – overlay GetCoords + GetEnv
output: summary statistics (mean, median, mode, quantiles, SD, range, etc.)


```{r}
library(dismo)
library(jsonlite)
library(tidyverse)

```



```{r}
#need to replace that in the input data
sstmean <- raster("bio oracle/Present.Surface.Temperature.Mean.asc")
sstmin <- raster("bio oracle/Present.Surface.Temperature.Min.asc")
sstmax <- raster("bio oracle/Present.Surface.Temperature.Max.asc")
sstrange <- raster("bio oracle/Present.Surface.Temperature.Range.asc")

#sp_data = name of the data frames with names of species and coordinate
lon<-sp_data$lon
lat<-sp_data$lat

sp.occr.latlon<-as.data.frame(cbind(lon,lat))

sp_meanT <- as.data.frame(extract(sstmean, sp.occr.latlon))
sp_minT <- as.data.frame(extract(sstmin, sp.occr.latlon))
sp_maxT <- as.data.frame(extract(sstmax, sp.occr.latlon))
sp_rangeT <- as.data.frame(extract(sstrange, sp.occr.latlon))

colnames(sp_meanT)<-"T_mean"
colnames(sp_minT)<-"T_min"
colnames(sp_maxT)<-"T_max"
colnames(sp_rangeT)<-"T_range"

sp_data<-cbind(sp_data,sp_meanT,sp_minT,sp_maxT,sp_rangeT)



```

