---
title: "GrossAff"
author: "Dvora"
date: "18 12 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---


GrossAff - Gross affinity from GBIF data 

input – overlay GetCoords + GetEnv
   
  1.  rasters - 'RasterStack' file, with layers of temperature. The choosen layres writen in "names" raw.(e.g BO_sstmax, BO_sstmean )

  2. data frame with species names, lat, lon

output: summary statistics (mean, median, mode, quantiles, SD, range, etc.)




Load packages
```{r}
library(summarytools)
library(tidyverse)
library(sdmpredictors)
library(dismo)
library(jsonlite)


```

cerate example data, with coordinate for Boops boops.
the data coords has the coordinate, with "lat" and "lon" column
`coords` should be a dataframe of coordinates (lat and lon)


```{r}
sp_data = read.csv(file = "sp.data.CSV")
  sp_data = sp_data %>% filter(species=="Boops boops")
    coords = sp_data %>% dplyr:: select(lon,lat)

```


overlap the coordinates and the temperatures layers from the raster
and create tible withe the summary for any layer

```{r}
list_of_layer_stats <- lapply(names(rasters),function(layer_name) {
  
  # extract according to 'coords'
temp <- raster::extract(rasters[[layer_name]], coords, df = TRUE)  dataframe (lon-lat)
  sumdata <- summarytools::descr(temp)
  sumdata = as_tibble(t(sumdata))[1,]
})

names(list_of_layer_stats) <- names(rasters)
  stat_dataframe <- bind_rows(list_of_layer_stats,.id = "id") %>% rename(layer = id)
```






summary of the temperature
summary statistics (mean, median, mode, quantiles, SD, range, etc.)

```{r}

```

