---
title: "TAFFY"
author: "Belmaker Lab"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(spocc)
library(sdmpredictors)
library(raster)
library(rfishbase)
library(summarytools)
library(dismo)
library(jsonlite)
library(mopa)
library(tidyverse)

```

1. **GetCoords**: Download data from [GBIF](https://www.gbif.org/) and [OBIS](https://obis.org/) | *Ori and Sarah*
   
   - download coordinate data by species
   - clean data [bdclean](https://bd-r.github.io/bdclean-guide/)
   - input: species list
   - output: coordinates dataframe
   
```{r}
# get coordinates from GBIF by species name

get_sp_coords <- function(species_name, site){
  if (site == "gbif"){
    sp=occ(species_name, "gbif", has_coords=TRUE)
    
    occs <- (sp$gbif$data[[1]][,2:3])
    
    occs$species <- species_name
    return(occs)
  }
}

get_sp_coords("Diplodus sargus", "gbif")

```

  
2. **FishBase data**: Load data from FishBase | *Shahar*

   - package [rfishbase](https://github.com/ropensci/rfishbase)
   - input: species names list
   - output: temperature range (preferred) + user requirements

This code uses a species list and binds it with its min, mean, and max preffered temperature from FishBase

```{r}
# Species list for the loop
species <- data$species_column

# Create an empty list to store the output
data_fbtemp_list <- list()

for(i in species) {
  # rfishbase function extracts min, mean and max preferred temperature for the species
  data_fbtemp_list[[i]] <- print(rfishbase::estimate(i) %>% select(contains("Temp")))
  }

# The output of the loop is a list so we'll unlist it into a dataframe:
data_temp <- data.frame(matrix(unlist(data_fbtemp_list), nrow=length(data_fbtemp_list), byrow=T))

# Rename columns
data_temp <- data_temp %>%
  rename(TempPrefMin = X1,
         TempPrefMean = X2,
         TempPrefMax = X3)

# Bind species list with temperature data
data_temp <- cbind(species,data_temp)

```

3. **GetEnv**: Load environmental data (e.g. temperature) from [Bio-ORACLE](http://www.bio-oracle.org/code.php) and [GMED](http://gmed.auckland.ac.nz/) | *Ori and Sarah*

   - input: Define a resolution and extent.
   - Load environmental layers

```{r}

rasters <- load_layers(c("BO_sstmax","BO_sstmean"))

# check raster resolution 
res(rasters)

# aggregate for increasing resulution of grid cell
aggregate(rasters, fact = 3)

# disaggregate for decreasing resulution of grid cell
disaggregate(rasters, fact = 4)

# fact = aggregation or disaggregation factor expressed as number of cells in each direction

```


4. **GrossAff** - Gross affinity from GBIF data | *Dvora*

   - input – overlay GetCoords + GetEnv
   - output: summary statistics (mean, median, mode, quantiles, SD, range, etc.)

Create example data, with coordinate for *Boops boops*:
`coords` = dataframe of coordinates (`lat` and `lon`)
> Note: All data has to be projected in WGS-84 `crs("+proj=longlat +datum=WGS84 +no_defs")`

```{r}
sp_data <- read_csv("sp.data.csv") %>%
  filter(species == "Boops boops") # Change the species name for the desired species

coords <- sp_data %>% dplyr::select(lon, lat)

```

Overlap coordinates and temperature layers from a raster and create tibble of layer's summary

```{r}
list_of_layer_stats <- lapply(names(rasters),function(layer_name) {
  # extract according to 'coords'
    temp <- raster::extract(rasters[[layer_name]], coords, df = TRUE) 
      sumdata <- summarytools::descr(temp)
      sumdata = as_tibble(t(sumdata))[1,]
  })

# join the names of the layers and convert the list of summaries to 1 data frame, "stat_dataframe"
names(list_of_layer_stats) <- names(rasters)
stat_dataframe <- bind_rows(list_of_layer_stats,.id = "id") %>%
  rename(layer = id) %>% 
  mutate(Range = Max - Min) # add range collumn

```

5. **GeoRange**: Estimate geographical range | *Yoni*

   - Create a grid by resolution from GetEnv
   - input: GetCoords, Fishbase data
   - Create convex hulls
   - Create alpha hulls (set smoothness) - Make sure n is sufficient; Overlay with land data and mask with depth

```{r}
# `coords` = from GetCoords stage

convex_hull <- convHull(coords)
polygon <- polygons(convex_hull)

# alpha hulls?

```

6. **GeoRangeAff**: Affinity by geographical range | *Shira*

   - input – GetEnv + GeoRange
   - output – summary statistics (mean, median, mode, quantiles, SD, range, Probabilities by degrees)

`rasters` = raster layers from `GetEnv` stage
`polygon` = convex halls converted to polygons from `GeoRange` stage

```{r}
list_of_layer_stats <- lapply(names(rasters), function(layer_name) {
  temp <- raster::extract(rasters[[layer_name]], polygon, df = TRUE) # extract raster values for polygon
  sumdata <- summarytools::descr(temp, stats = "common", transpose = TRUE) # summary statistics for  temperature data
  sumdata = as_tibble(sumdata)[1,]
})

names(list_of_layer_stats) <- names(rasters) # names of layers
stat_dataframe <- bind_rows(list_of_layer_stats,.id = "id") %>% rename(layer = id) # bind all together and create a summary statistics dataframe


```

7. **SDMaff** Affinity by SDM | *Itai G. and Yoni*

   - Define pseudoabsences ([mopa package](https://www.rdocumentation.org/packages/mopa/versions/1.0.1))
   - Spatial thining
   - Create SDM (most recent package)
   - Summary statistics (mean, median, mode, quantiles, SD, range, probability by degrees)

Funnctions to perform SDMs and calculate species level thermal affinity matrices 
Major input: GetCoords + GetEnv
Coords - data frame of z/ y (long / lat). Output of GetCoords
Env - stacked enviromental layer. Output of GetEnv
```{r}
# Define pseudoabsences, uses the the novel three-step method as described in Iturbide et al. (2015) 
 
GetPseudo <- function(Coords, Env, N, prev) {
  bg.profile <- OCSVMprofiling(xy = Coords, varstack = Env)
  bagr.radius <- backgroundRadius(xy = Coords, background = bg.profile$absence,  unit = "decimal degrees") 
  pseudo_TS <- pseudoAbsences(xy = Coords, realizations=N, background = bagr.radius, exclusion.buffer = 0.0166, prevalence = prev)
  return(pseudo_TS) 
}

# Perform SDM
# Note that this SDM does not perform cross-validation

SDM <- function(Coords, Env, Pseudo, real) {
  library(dismo)
  return(maxent(p=Coords, x=Env, a=Pseudo[[real]]))
}

```

Extract affinity
`temp` = enviromental variable to extract, usualy temperature.

The output is currently a vector of the Quantiles = c(0.1,0.25,0.50,0.75,0.9)as well as optimal values for the selcte enviroemntal predictor 
```{r}
SDMaff <- function(SDM, temp){
  a <- as.data.frame(response(SDM,var = temp))
  a$prob <- a$p/sum(a$p)
  a$cum <- cumsum(a$prob)
  Quant <- c(0.1,0.25,0.50,0.75,0.9)
  Locations <- sapply(Quant, function(y)which.min(abs(a[,4]-y)))
  Topt <- a[which.min(abs(max(a$p)-a$p)),1]
  Tq10 <- a[Locations[1],1]
  Tq25 <- a[Locations[2],1]
  Tq50 <- a[Locations[3],1]
  Tq75 <- a[Locations[4],1]
  Tq90 <- a[Locations[5],1]
  
  Values <- cbind.data.frame(c("Topt","Tq10","Tq25","Tq50","Tq75","Tq90"), c(Topt,Tq10,Tq25,Tq50,Tq75,Tq90))
  
  return(Values)
}  

```

**The  actual code to run!!!!**
```{r}
N <- 5 # N = number of pseudoAbsences realizations
prev <- 0.5 # ratio of presences to pseudoAbsences
Affinity <- matrix(NA, nrow = N, ncol = 6)
colnames(Affinity) <- c("Topt","Tq10","Tq25","Tq50","Tq75","Tq90")

# temp = enviromental variable to extract, usualy temperature

for (real in 1:N ) {  # loop to go over differnt pseudoAbsences realizations  
  
  pseudo_TS<- GetPseudo(Coords, Env, N, prev)
  SDM_output <- SDM(Coords, Env, Pseudo, real) 
  Affinity[real, ]<-SDMaff(SDM,temp) 
  
} # end of loop to go over differnt pseudoAbsences realizations

FinalAffinity <- colMeans(Affinity) # the final climatic affintiy estimates 

```

8. **Cheung affinity** | *Dvora and Daphna*
   – From [Cheung's paper](https://www.nature.com/articles/nature12156)
This function is getting Cheung Temperature affinity. He create table of 991 species from the marine ecosystem (not fish only).
input:`species` - species name, vector or dataframe, could recieved 1 species name or a vector/data frame with a lot of species names.
output: `species_tp` - data frame with TP50, TP75, TP25 - temperature preference of the species, if it exist in Chueng's table (NA if not).

Note: If you can't retrieve temperature - try a different species name (check in [WoRMS](http://www.marinespecies.org/))
```{r}
# upload Chueng data
ch_data <- read_csv("Chueng data clean.csv")

# replace the name of species column, and remove common names
ch_data <- ch_data %>% rename(species = Latin) %>% select(-English)

# convert TP columns to numeric
ch_data$TP25 <- as.numeric(ch_data$TP25)
ch_data$TP75 <- as.numeric(ch_data$TP75)

# left join
species_tp <- as.data.frame(species)
species_tp <- species_tp %>% left_join(ch_data)


```



```

